---
title: "HW2-exploration"
format: html
editor: visual
---

## HW2 Exploration

Data derived from paper titled: Contrasting effects of vineyard type, soil and landscape factors on ground- versus above-ground nesting bees

https://datadryad.org/stash/dataset/doi:10.5061/dryad.h44j0zpq0

Citation: Wersebeckmann, Vera; Warzecha, Daniela; Entling, Martin H.; Ilona, Leyer (2023). Contrasting effects of vineyard type, soil and landscape factors on ground- versus above-ground nesting bees \[Dataset\]. Dryad. <https://doi.org/10.5061/dryad.h44j0zpq0>

```{r}
#..........................load packages.........................
library(tidyverse)
library(chron)
library(naniar)
library(here)

# read in data
# use read_csv2() for files with a separator of ;
bees <- read_csv2(here::here("data", "doi_10_5061_dryad_h44j0zpq0__v20221222", "Wersebeckmann_et_al_wildbee_species_matrix_allSR.csv"))
                
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
